<!-- LTeX: language=en-US -->
# $\mathrm{p}$-Values: Challenges for Hypothesis Testing


#### 🎯 **Lecture Goals** {-}


* **How** do $\mathrm{p}$-values work?

  * Today: [two-sided $t$-test]{style="background-color: yellow"}

* **Why** are $\mathrm{p}$-values a challenge for hypothesis testing?

* **What** alternatives are there? 

::: {.hidden}
$$
\require{color}
%% Colorbox within equation-environments:
\newcommand{\highlight}[2][yellow]{\mathchoice%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  }%
$$
:::


```{r}
#| echo: false
library("scales")
library("extraDistr")

value <- 0
value <- value + 1
set.seed(33)
n     <- 5
Y_1   <- round(rnorm(n, mean = 5), 2)
Y_2   <- round(rnorm(n, mean = 5), 2)
X     <- Y_2 - Y_1
X_bar <- round(mean(X), 1)
s     <- round(sd(X), 1)
T_obs <- round((sqrt(n) * X_bar)/s, 2)
```


## Real-Example: Marketing Campaign


Marketing campaign in $n=5$ stores (paired sample): 

</br>

Store | Sales Before | Sales After | Difference (After $-$ Before)
----|------------|------------|--------- 
$1$   |$`r format(Y_1[1], nsmall=2)`$|$`r format(Y_2[1], nsmall=2)`$|$X_{1,obs}=`r format(X[1], nsmall=2)`$
$2$   |$`r format(Y_1[2], nsmall=2)`$|$`r format(Y_2[2], nsmall=2)`$|$X_{2,obs}=`r format(X[2], nsmall=2)`$
$3$   |$`r format(Y_1[3], nsmall=2)`$|$`r format(Y_2[3], nsmall=2)`$|$X_{3,obs}=`r format(X[3], nsmall=2)`$
$4$   |$`r format(Y_1[4], nsmall=2)`$|$`r format(Y_2[4], nsmall=2)`$|$X_{4,obs}=`r format(X[4], nsmall=2)`$
$5$   |$`r format(Y_1[5], nsmall=2)`$|$`r format(Y_2[5], nsmall=2)`$|$X_{5,obs}=`r format(X[5], nsmall=2)`$  

</br>


@fig-diffData shows the observed sales differences 
$$
\{X_{1,obs},X_{2,obs},X_{3,obs},X_{4,obs},X_{5,obs}\}.
$$ 

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 2
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-diffData
#| fig-cap: "Sales differences (sales after minus before the marketing campaign)."
par(mar = c(0, 0, 0, 0))
plot(x = X, y = rep(0, n), 
     pch = 21, 
     col = alpha("red", 0.5), 
     bg  = alpha("red", 0.5), 
     cex = 1.5, axes = FALSE, 
     ylab = "", xlab = "", xlim = c(-3,3))
axis(1, at = -2:2, line = -3.5, cex.axis = 1.5, lwd = 2, family="serif")
```

::: {.callout-note appearance="minimal"}
**Assumption:** There are no systematic differences (mean-shifts) between sales before and after the marketing campaign, other than the potential impact of the campaign itself.
:::

</br>


* **Research Question:**  </br>Was the marketing campaign successful? Did it show a tendency (location-shift) toward *significantly* higher or lower sales?

* **Statistical Reformulation:** </br>Is the mean of $X_1,\dots,X_n$ different from zero?

* **Testing problem:** $\hspace{1cm}H_0\colon \mu = 0\quad$ versus $\quad H_1\colon \mu \neq 0$

<!-- * **Setup:**</br>Random sample $X_1,\dots,X_n\overset{iid}{\sim}X$ with $\mu_X = E(X)$ -->



## Short Repetition: Two-Sided $t$-Test

::: {.callout-note appearance="minimal"}
**More Details:** See our lecture on hypothesis testing for more details. 
:::



Statistical hypothesis:

<center>
$\displaystyle H_0\colon \mu = 0\quad$ versus $\displaystyle \quad H_1\colon \mu \neq 0$
</center>

</br>

$t$-test statistic:
$$
\mathrm{T} = \frac{\sqrt{n}(\bar{X} - 0)}{s}\overset{H_0}{\sim} t_{n-1},
$$
where 
\begin{align*}
\bar{X}&=\frac{1}{n}\sum_{i=1}^n X_i\\
s      &= \sqrt{\frac{1}{n-1}\sum_{i=1}^n(X_i-\bar{X})}
\end{align*}

denote the empirical mean and the empirical standard deviation.

::: {.callout-note appearance="minimal"}
**Assumptions:**

* Random (iid) sample: 
$X_1,\dots,X_n\overset{\operatorname{iid}}{\sim}X$ 

* For small sample sizes: 
$X\sim\mathcal{N}(\mu, \sigma^2)$ 

* For large sample sizes (CLT, $n\to\infty$): 
$$
\mathrm{E}(X)=\mu\quad\text{and}\quad \mathrm{Var}(X)=\sigma^2
$$
:::

</br>

**Observed (*obs*) realization (marketing campaign):**

</br>

$$
\mathrm{T}_{\mathrm{obs}} = \frac{\sqrt{n}(\bar{X}_{\mathrm{obs}} - 0)}{s_{\mathrm{obs}}} = `r format(T_obs, nsmall=2)`
$$ 

* Empirical mean: $\bar{X}_{\mathrm{obs}}=\frac{1}{5}\sum_{i=1}^5 X_{i,obs} = `r format(X_bar,nsmall=2)`$ 
* Empirical standard deviation: $s_{\mathrm{obs}} = \sqrt{\frac{1}{5-1}\sum_{i=1}^5(X_{i,obs}-\bar{X}_{\mathrm{obs}})} = `r format(s,nsmall=2)`$ 



#### **Critical Values $\phantom{\mathrm{p}}$** {-}

::: {.callout-note appearance="minimal"}
**Critical Values:** The critical values — the $(1-\alpha/2)$- and the $\alpha/2$-quantiles of the $t$-distribution with $\operatorname{df}=n-1$ — define the rejection regions for a given significance level $0<\alpha<1$.
:::

Marketing campaign data:

::: {.r-stack}
Density function of the $t_{\operatorname{df}}$ distribution with $\operatorname{df}=4$
:::

```{r}
#| echo: false
#| fig-width: 7
#| out-width: 7in
#| fig-align: center

alpha  <- 0.05
df     <- 5-1
q_1    <- qt(p = alpha/2,   df=df)
q_2    <- qt(p = 1-alpha/2, df=df)

xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)

par(mar = c(4.5, .1, .1, .1))
curve(expr = dt(x = x, df = df),
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      ylim = c(0, .4), xaxs="i",yaxs="i", axes = FALSE)

polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = alpha("red", 0.25))
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = alpha("red", 0.25))
axis(1, cex.axis = 2, family = "serif")
axis(1, at = T_obs, col = "blue", labels = FALSE, tck=-0.11)
axis(1, at = q_1,   col = "red",  labels = FALSE, tck=-0.11)
lines(x = c(q_1,q_1), y = c(0, dt(x = q_1, df = df)), col="red")
axis(1, at = q_2,   col = "red",  labels = FALSE, tck=-0.11)
lines(x = c(q_2,q_2), y = c(0, dt(x = q_2, df = df)), col="red")
#mtext(side = 1, text = c(expression(q[alpha/2]), bquote(T[obs]==.(T_obs)), expression(q[1-alpha/2])), at = c(q_1,T_obs,q_2), line = 3.5, cex = 2, family = "serif")
mtext(side = 1, text = c(expression(q[alpha/2]), expression(T[obs]), expression(q[1-alpha/2])), at = c(q_1,T_obs,q_2), line = 3.5, cex = 2, family = "serif")
legend("topright", pch=c(22), cex=2.1, 
       col = alpha("red", 0.25), 
       legend=expression(alpha==0.05), 
       bty="n", pt.bg = alpha("red", 0.25))
```


 

## Two-Sided $\mathrm{p}$-Value

::: {.callout-note appearance="minimal"}
**$\mathrm{p}$-Value:** The $\mathrm{p}$-value is the probability of observing a test statistic $\mathrm{T}$ as extreme as $\mathrm{T}_{\mathrm{obs}}$ or more extreme (in the direction(s) of the alternative hypothesis) than $\mathrm{T}_{\mathrm{obs}}$, given the null hypothesis ($H_0$) is true.

* **Two-Sided $\mathrm{p}$-Value:** 
$$\displaystyle\mathrm{p}_{\mathrm{obs}} = \mathrm{P}\big(\,|\mathrm{T}|\,\geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big)$$
:::



Marketing campaign data:


::: {.r-stack}
Density function of the $t_{\operatorname{df}}$ distribution with $\operatorname{df}=4$
:::

```{r}
#| fig-width: 7
#| out-width: 7in
#| fig-align: center

alpha  <- 0.05
df     <- 5-1
q_1    <- -abs(T_obs)
q_2    <- abs(T_obs)

pValue <- round(2 * min(pt(q_1, df = df), pt(q_2, df = df)), 2)

xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)

par(mar = c(4.5, .1, .1, .1))
curve(expr = dt(x = x, df = df),
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      ylim = c(0, .4), xaxs="i",yaxs="i", axes = FALSE)

polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = alpha("red", 0.25))
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = alpha("red", 0.25))
axis(1, cex.axis = 2, family = "serif")
axis(1, at =  T_obs, col = "red", labels = FALSE, tck=-0.11)
axis(1, at = -T_obs, col = "red", labels = FALSE, tck=-0.11)
mtext(side = 1, text = expression(-T[obs], T[obs]), at = c(-abs(T_obs),abs(T_obs)), line = 3.5, cex = 2, family = "serif")
legend("topright", pch=c(22), cex=2.1, 
       col = alpha("red", 0.25), 
       legend=bquote("p-value"==.(pValue)), 
       bty="n", pt.bg = alpha("red", 0.25))
```



## Two-Sided $\mathrm{p}$-Value

::: {.callout-note appearance="minimal"}
**$\mathrm{p}$-Value:** The $\mathrm{p}$-value is the probability of observing a test statistic $\mathrm{T}$ as extreme as $\mathrm{T}_{\mathrm{obs}}$ or more extreme (in the direction(s) of the alternative hypothesis) than $\mathrm{T}_{\mathrm{obs}}$, given the null hypothesis ($H_0$) is true.

* **Two-Sided $\mathrm{p}$-Value:** 
$$\displaystyle\mathrm{p}_{\mathrm{obs}} = \mathrm{P}\big(\highlight{|\mathrm{T}|} \geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big)$$
:::

Marketing campaign data:


::: {.r-stack}
Density function of the [Half-$t_{\operatorname{df}}$]{style="background-color: yellow"} distribution with $\operatorname{df}=4$
:::

```{r}
#| fig-width: 7
#| out-width: 7in
#| fig-align: center

library("extraDistr")

alpha  <- 0.05
df     <- 5-1
q_1    <- -abs(T_obs)
q_2    <- abs(T_obs)

pValue <- round(2 * min(pt(q_1, df = df), pt(q_2, df = df)), 2)

xx0    <- seq(0.00001, 5, len=50)
##
xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)
##
xx3    <- seq(abs(T_obs),  5, len=25)
yy3    <- dht(x = xx3, nu = df)

par(mar = c(4.5, .1, .1, .1))
plot(y = dht(x = xx0, nu = df), x = xx0,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      xaxs="i",yaxs="i", axes = FALSE, type="l")
lines(x=c(0,0), y = c(0,dht(x = xx0, nu = df)[1]), lwd=2)
curve(expr = dt(x = x, df = df), 
      lwd = 2, col = gray(.75), add = TRUE)


polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = gray(.95))
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = gray(.95))
polygon(x = c(xx3,rev(xx3)), y=c(yy3, rep(0,length(yy3))), border = NA, col = alpha("red", 0.25))
axis(1, cex.axis = 2, family = "serif")
axis(1, at =  T_obs, col = "red", labels = FALSE, tck=-0.11)
axis(1, at = -T_obs, col = gray(.75), labels = FALSE, tck=-0.11)
mtext(side = 1, text = expression("|"~T[obs]~"|"), at = c(abs(T_obs)), line = 3.5, cex = 2, family = "serif")
legend("topright", pch=c(22), cex=2.1, 
       col = alpha("red", 0.25), 
       legend=bquote("p-value"==.(pValue)), 
       bty="n", pt.bg = alpha("red", 0.25))
```



## Two-Sided $\mathrm{p}$-Value

<!-- 
::: {.callout-note appearance="minimal"}
**$\mathrm{p}$-Value:** The $\mathrm{p}$-value is the probability, assuming the null hypothesis ($H_0$) is true, of observing a test statistic $\mathrm{T}$ as extreme as $\mathrm{T}_{\mathrm{obs}}$ or more extreme (in the direction(s) of the alternative hypothesis) than $\mathrm{T}_{\mathrm{obs}}$. 
::: 

::: {.callout-note appearance="minimal"}
Thus for two-sided test statistics, we have
\begin{align*}
\mathrm{p}_{\mathrm{obs}} 
& = \mathrm{P}\big(|\mathrm{T}| \geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big).
\end{align*}
:::
-->

:::: {.columns}

::: {.column width="50%"}
**Two-sided $t$-test:** 
<!-- 
Since $\mathrm{T}\overset{H_0}{\sim} t_{n-1}$, we have that [$|\mathrm{T}|\overset{H_0}{\sim} \text{Half-}t_{n-1}$]{style="background-color: yellow"} 
-->
\begin{align*}
\mathrm{p}_{\mathrm{obs}} 
& = \mathrm{P}\big(\,|\mathrm{T}|\,\geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big)\\[1ex]
& = 1 - \mathrm{P}\big(\highlight{|\mathrm{T}|}\leq |\mathrm{T}_{\mathrm{obs}}|\;\big|\highlight{H_0\;\text{is true}}\big)\\[1ex]
& = 1 - \highlight{\mathrm{F}_{\,\text{Half-}t,n-1}}\big(|\mathrm{T}_{\mathrm{obs}}|\big),
\end{align*}
where 
$\mathrm{F}_{\,\text{Half-}t,n-1}$ denotes the cdf of the Half-$t$ distribution with $\operatorname{df}=n-1$
:::
::: {.column width="50%"}

```{r}
#| fig-width: 7
#| out-width: 7in
#| fig-align: center

alpha  <- 0.05
df     <- 5-1

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5))
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, 1), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.75, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.75, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_obs, col = "blue", labels = FALSE, tck=-0.11)
axis(2, at =  myfun(T_obs), col = "blue", labels = FALSE, tck=-0.05)
lines(x = c(0,T_obs), y= rep(myfun(T_obs), 2), col="blue")
lines(x = rep(T_obs, 2), y= c(0, myfun(T_obs)), col="blue")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"), at = c(abs(T_obs)), line = 3.5, cex = 2, family = "serif", hadj = 1)
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_obs)), bquote(p[obs]), srt = 0, cex = 2, family = "serif")


legend("top", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```
:::
::::

```{r}
#| echo: true
library("extraDistr")
df <- 5-1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)
```

::: {.notes}
* There's a one-to-one mapping between $|\mathrm{T}_{obs}|$ and $p_{\mathrm{obs}}$. 
* $p_{\mathrm{obs}}$ effectively measures how extrem the value of $|\mathrm{T}_{obs}|$ is, given that $H_0$ is true.
:::


## Two-Sided $\mathrm{p}$-Value


:::: {.columns}

::: {.column width="50%"}
Marketing campaign data:

* $|\mathrm{T}_\mathrm{obs}|=`r T_obs`$
* $\alpha = 0.05$
* $(1-\alpha/2)$-quantile of $t_{4}\colon$</br> $\mathrm{q}_{1-\alpha/2}=`r round(qt(p = 1-alpha/2, df=df),2)`$ 
:::
::: {.column width="50%"}
```{r}
#| fig-width: 7
#| out-width: 7in
#| fig-align: center

alpha  <- 0.05
df     <- 5-1

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5))
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, 1), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.75, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.75, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_obs, col = "blue", labels = FALSE, tck=-0.11)
axis(2, at =  myfun(T_obs), col = "blue", labels = FALSE, tck=-0.05)
axis(1, at =  qt(1-alpha/2, df=df), col = "red", labels = FALSE, tck=-0.1)
axis(2, at = alpha, col = "red", labels = FALSE, tck=-0.05)
lines(x = c(0,T_obs), y= rep(myfun(T_obs), 2), col="blue")
lines(x = rep(T_obs, 2), y= c(0, myfun(T_obs)), col="blue")
lines(x = c(0,qt(1-alpha/2, df=df)), y= rep(alpha, 2), col="red")
#lines(x = c(0, 0), y= c(0, alpha), lwd = 3, col = "red")
lines(x = rep(qt(1-alpha/2, df=df), 2), y= c(0, alpha), lwd = 1, col = "red")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"), at = c(abs(T_obs)), line = 3.5, cex = 2, family = "serif", hadj = 1)
mtext(side = 1, text = bquote(q[1-alpha/2]), at = qt(1-alpha/2, df=df), line = 3.5, cex = 2, family = "serif", hadj = 1)
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_obs)), bquote(p[obs]==.(round(myfun(T_obs), 2))), srt = 0, cex = 2, family = "serif")
text(x = corners[1]-.75, y = alpha, bquote(alpha==.(alpha)), srt = 0, cex = 2, family = "serif")

legend("top", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```
:::
::::
<!-- 
```{r}
#| echo: true
library("extraDistr")
df <- 5-1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)
``` 
-->


We fail to reject $H_0\colon\mu = 0$. That is, we have no statistical evidence that the marketing campaign showed a tendency toward *significantly* higher or lower sales. 



::: {.callout-caution appearance="minimal"}
⚠️ Failing to reject $H_0$ does not mean that $H_0$ is true. It only means that we do not have enough evidence to reject it.
:::


## Two-Sided $\mathrm{p}$-Value


[Equivalent test decisions]{style="background-color: yellow"} at the significance level $0<\alpha<1$: 


:::: {.columns}

::: {.column width="45%"}
* **Critical values:**</br> 
Reject $H_0\colon \mu =0$ if [$\displaystyle|\mathrm{T}_{\mathrm{obs}}| > q_{1-\alpha/2}$]{style="background-color: yellow"}, 
   where $q_{1-\alpha/2}$ denotes the $(1-\alpha/2)$-quantile of the $t_{\operatorname{df}}$ distribution. 

</br>

* **$\mathrm{p}$-Values:**</br>
Reject $H_0\colon \mu =0$ if [$\displaystyle\mathrm{p}_{\mathrm{obs}} < \alpha$]{style="background-color: yellow"} 
:::
::: {.column width="55%"}
```{r}
#| fig-width: 7
#| out-width: 7in
#| fig-align: center

alpha  <- 0.05
df     <- 5-1

T_2obs <- 3.7

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5))
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, .2), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.75, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.75, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_2obs, col = "blue", labels = FALSE, tck=-0.11)
axis(2, at =  myfun(T_2obs), col = "blue", labels = FALSE, tck=-0.05)
axis(1, at =  qt(1-alpha/2, df=df), col = "red", labels = FALSE, tck=-0.1)
axis(2, at = alpha, col = "red", labels = FALSE, tck=-0.05)
lines(x = c(0,T_2obs), y= rep(myfun(T_2obs), 2), col="blue")
lines(x = rep(T_2obs, 2), y= c(0, myfun(T_2obs)), col="blue")
lines(x = c(0,qt(1-alpha/2, df=df)), y= rep(alpha, 2), col="red")
#lines(x = c(0, 0), y= c(0, alpha), lwd = 3, col = "red")
lines(x = rep(qt(1-alpha/2, df=df), 2), y= c(0, alpha), lwd = 1, col = "red")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"), at = c(abs(T_2obs)), line = 3.5, cex = 2, family = "serif", hadj = 1)
mtext(side = 1, text = bquote(q[1-alpha/2]), at = qt(1-alpha/2, df=df), line = 3.5, cex = 2, family = "serif", hadj = 1)
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_2obs)), bquote(p[obs]), srt = 0, cex = 2, family = "serif")
text(x = corners[1]-.75, y = alpha, bquote(alpha==.(alpha)), srt = 0, cex = 2, family = "serif")


legend("topright", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```
:::
::::



## $\mathrm{p}$-Values: Challenges for Hypothesis Testing

* While the $\mathrm{p}$-value can be a useful statistical measure, it is commonly misused and misinterpreted. (@Wasserstein2016asa, @wasserstein2019asa)


* "Scientific method: Statistical errors." @nuzzo2014scientific

* ScienceNews (Siegfried 2010): "It’s science's dirtiest secret: The 'scien-
tific method' of testing hypotheses by statistical analysis stands on a flimsy foundation."


## References

::: {#refs}
:::


## 💡 Tips for Success

<br>

* Don’t just memorize formulas — focus on understanding *why* they work.
* Discuss and collaborate

