<!-- LTeX: language=en-US -->
# $\mathrm{p}$-Values: Challenges for Hypothesis Testing



::: {.content-hidden unless-format="html"}
$$
\require{color}
%% Colorbox within equation-environments:
\newcommand{\highlight}[2][yellow]{\mathchoice%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  }%
$$
:::

::: {.content-hidden unless-format="pdf"}
```{=latex}
%% Colorbox within equation-environments:
\newcommand{\highlight}[2][yellow]{\mathchoice%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  }%
```
:::

## Introduction 

### 🎯 Lecture Goals {-}

In this lecture, we’ll tackle three key questions:

* **How** do $\mathrm{p}$-values actually work?
* **Why** can $\mathrm{p}$-values be tricky or even misleading in hypothesis testing?
* **What** is good practice when working with $\mathrm{p}$-values?

We’ll explore these questions using the **two-sided $t$-test** as our main example. Parallel arguments apply to one-sided testing and other hypothesis tests (F-test, etc.)

```{r}
#| echo: false
library("scales")
library("extraDistr")

value <- 0
value <- value + 1
set.seed(33)
n     <- 5
Y_1   <- round(rnorm(n, mean = 5), 2)
Y_2   <- round(rnorm(n, mean = 5), 2)
X     <- round(Y_2 - Y_1, 2)
X_bar <- round(mean(X), 2)
s     <- round(sd(X),   2)
T_obs <- round((sqrt(n) * X_bar)/s, 2)
```


### 💻 Example: Marketing Campaign {-}

Imagine a company runs a marketing campaign in $n = 5$ stores and measures sales *before* and *after* the campaign.

</br>

Store | Sales Before | Sales After | Difference (After $-$ Before)
----|------------|------------|--------- 
$1$   |$`r format(Y_1[1], nsmall=2)`$|$`r format(Y_2[1], nsmall=2)`$|$X_{1,\mathrm{obs}}=`r format(X[1], nsmall=2)`$
$2$   |$`r format(Y_1[2], nsmall=2)`$|$`r format(Y_2[2], nsmall=2)`$|$X_{2,\mathrm{obs}}=`r format(X[2], nsmall=2)`$
$3$   |$`r format(Y_1[3], nsmall=2)`$|$`r format(Y_2[3], nsmall=2)`$|$X_{3,\mathrm{obs}}=`r format(X[3], nsmall=2)`$
$4$   |$`r format(Y_1[4], nsmall=2)`$|$`r format(Y_2[4], nsmall=2)`$|$X_{4,\mathrm{obs}}=`r format(X[4], nsmall=2)`$
$5$   |$`r format(Y_1[5], nsmall=2)`$|$`r format(Y_2[5], nsmall=2)`$|$X_{5,\mathrm{obs}}=`r format(X[5], nsmall=2)`$  

</br>


@fig-diffData visualizes the observed differences $X_{i,\mathrm{obs}},$ $i=1,\dots,5.$
<!-- \begin{align*}
&\{X_{1,obs},X_{2,obs},X_{3,obs},X_{4,obs},X_{5,obs}\}\\
=&\{`r format(X[1], nsmall=2)`, `r format(X[2], nsmall=2)`, `r format(X[3], nsmall=2)`, `r format(X[4], nsmall=2)`, `r format(X[5], nsmall=2)`\}
\end{align*} -->

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 0.75
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-diffData
#| fig-cap: "Sales differences (sales after minus before the marketing campaign)."
par(mar = c(2, 0, 0, 0))
plot(x = X, y = rep(0.002, n), 
     pch = 21, 
     col = alpha("red", 0.5), 
     bg  = alpha("red", 0.5), 
     cex = 2, axes = FALSE, 
     ylab = "", xlab = "", xlim = c(-2.1,2.1), ylim = c(0,.005), yaxs="i")
axis(1, at = -2:2, cex.axis = 1.5, lwd = 2, family="serif", pos =0)
```


**We want to know:**

* **Research question:**  </br> 
Was the marketing campaign successful? Did it show a tendency (location-shift) toward *significantly* higher or lower sales?

* **Statistical question:** </br> 
Is the mean, $\mu,$ of $X_1,\dots,X_n$ different from zero?

<!-- * **Formal test:** $\hspace{1cm}H_0\colon \mu = 0\quad$ versus $\quad H_1\colon \mu \neq 0$ -->

::: {.callout-tip appearance="minimal"}
**Assumption:** There are no systematic differences (mean-shifts) between sales before and after the marketing campaign, other than the potential impact of the campaign itself.
:::

<!-- 
### 📖 Quick Recap: Two-Sided $t$-Test, Critical Values, and Decision {-} 

::: {.callout-tip appearance="minimal"}
For more background, see our earlier lecture on hypothesis testing.
:::

#### **Refresher: Two-Sided $t$-Test** {-}
-->

#### **Two-Sided $t$-Test** {-}
We test

<center>
$\displaystyle H_0\colon \mu = 0\quad$ versus $\displaystyle \quad H_1\colon \mu \neq 0$
</center>

</br>

The $t$-statistic is:
$$
\mathrm{T} = \frac{\bar{X} - 0}{S/\sqrt{n}}\overset{H_0}{\sim} t_{n-1},
$$
where 

* $\bar{X}=\frac{1}{n}\sum_{i=1}^n X_i$ is the **sample mean**, and
* $S= \sqrt{\frac{1}{n-1}\sum_{i=1}^n(X_i-\bar{X})}$ is the **sample standard deviation.**

::: {.callout-tip appearance="minimal"}
**Assumptions:**

* Random (iid) sample: 
$X_1,\dots,X_n\overset{\operatorname{iid}}{\sim}X$ 
* For small $n$: 
$X\sim\mathcal{N}(\mu, \sigma^2)$ 

* For large $n$: we just need finite mean and variance (Central Limit Theorem helps).
:::

</br>

#### **Observed (obs) $t$-Statistic (Marketing Campaign Data)** {-}

$$
\mathrm{T}_{\mathrm{obs}} = \frac{\bar{X}_{\mathrm{obs}} - 0}{S_{\mathrm{obs}}/\sqrt{5}} = `r format(T_obs, nsmall=2)`
$$ 

* $\bar{X}_{\mathrm{obs}}=\frac{1}{5}\sum_{i=1}^5 X_{i,\mathrm{obs}} = `r format(X_bar,nsmall=2)`$ 
* $S_{\mathrm{obs}} = \sqrt{\frac{1}{5-1}\sum_{i=1}^5(X_{i,\mathrm{obs}}-\bar{X}_{\mathrm{obs}})} = `r format(s,nsmall=2)`$ 
* Significance level: $\alpha=0.05$
* Critical values (quantiles of $t_{n-1}$): $\mathrm{q}_{\alpha/2} = `r round(qt(p = 0.05/2, df=4), 2)`\;$ and $\;\mathrm{q}_{1-\alpha/2} = `r round(qt(p = 1-0.05/2, df=4), 2)`$

<!-- 
This observed $t$-statistic value tells us how far the mean difference is from zero, measured in units of the estimated standard error. 
-->

```{r}
#| echo: true
#| eval: false
n     <- 5
X_obs <- c(0.64, -0.72, -0.23, 0.91, 1.06)
X_bar <- mean(X_obs)
S_obs <- sd(X_obs)

## Observed t-statistic
T_obs <- (X_bar - 0)/(S_obs/sqrt(n))

## Alternatively: 
result <- t.test(X_obs, mu = 0, alternative = "two.sided")
T_obs  <- result$statistic

## critical values
alpha <- 0.05
qt(p =     alpha/2, df = n-1)
qt(p = 1 - alpha/2, df = n-1)
```

<!-- 
#### **Refresher: Critical Values** {-}

::: {.callout-tip appearance="minimal"}
**Critical values** mark the "cutoff points" for deciding whether our observed statistic $\mathrm{T}_{\mathrm{obs}}$ is too extreme to be consistent with $H_0$.
:::

In case of the two-sided $t$-test, the **critical values** are the $(\alpha/2)$- and $(1 - \alpha/2)$-quantiles
<center>
$\displaystyle \mathrm{q}_{\alpha/2}\quad$ and $\displaystyle \quad\mathrm{q}_{1-\alpha/2}\quad$
</center>
of the $t$-distribution with $n-1$ degrees of freedom, where $0<\alpha<1$ denotes the significance level. @fig-CriticalValues shows the null distribution (the $t_{n-1}$ density) and the rejection regions for $\alpha = 0.05$. 
-->


#### **Test Decision (Marketing Campaign Data)** {-}

Since $\textrm{T}_{\textrm{obs}}$ is not in the rejection regions (see @fig-CriticalValues)
\begin{align*}
\textrm{T}_{\textrm{obs}}=`r T_obs` &\;\not\in\;(-\infty, \mathrm{q}_{\alpha/2})\; \cup \;(\mathrm{q}_{1-\alpha/2},\infty\,)
%\textrm{T}_{\textrm{obs}}=`r T_obs` &\;\not\in\;]\,\mathrm{q}_{1-\alpha/2},\infty\,[,
\end{align*}
we [fail to reject $H_0\colon\mu = 0$]{style="background-color: yellow; padding: 0.1em 0.1em;"}  at the significance level [$\alpha = 0.05.$]{style="background-color: #F88379CC; padding: 0.1em 0.1em;"}


```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-CriticalValues
#| fig-cap: "Density of the null-distribution $t_{\\operatorname{df}}$ with $\\operatorname{df}=n-1=4$ (marketing campaign data). The rejection regions are defined by the quantiles (critical values) $\\mathrm{q}_{\\alpha/2}$ and $\\mathrm{q}_{1-\\alpha/2}$ of $t_{\\operatorname{df}}$ for $\\alpha=0.05.$"


alpha  <- 0.05
df     <- 5-1
q_1    <- qt(p = alpha/2,   df=df)
q_2    <- qt(p = 1-alpha/2, df=df)

xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)

myRed  <- "#F88379CC"

par(mar = c(4.5, .1, .1, .1))
curve(expr = dt(x = x, df = df),
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      ylim = c(0, .4), xaxs="i",yaxs="i", axes = FALSE)

polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = myRed)
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = myRed)
curve(expr = dt(x = x, df = df),
      lwd = 2, col = 1, add = TRUE)
axis(1, cex.axis = 1.5, family = "serif")
axis(1, at = T_obs, col = "blue", labels = FALSE, tck=-0.09)
axis(1, at = q_1,   col = "red",  labels = FALSE, tck=-0.09)
lines(x = c(q_1,q_1), y = c(0, dt(x = q_1, df = df)), col="red")
axis(1, at = q_2,   col = "red",  labels = FALSE, tck=-0.09)
lines(x = c(q_2,q_2), y = c(0, dt(x = q_2, df = df)), col="red")
mtext(side = 1, text = c(expression(q[alpha/2]), expression(T[obs]), expression(q[1-alpha/2])), at = c(q_1,T_obs,q_2), line = 3.5, cex = 1.5, family = "serif")
legend("topleft", pch=c(22), cex=1.5, 
       col = myRed, 
       legend=expression(alpha==0.05), 
       bty="n", pt.bg = myRed)
```


## The Two-Sided $\mathrm{p}$-Value of the $t$-Test 

### Definition of the $\mathrm{p}$-Value

::: {.callout-tip appearance="minimal"}
## Definition of the $\mathrm{p}$-Value 
The $\mathrm{p}$-value, $\textrm{p}_{\mathrm{obs}},$ is the probability of observing a test statistic $\mathrm{T}$ as extreme as or more extreme than $\mathrm{T}_{\mathrm{obs}}$ (in the direction(s) of the alternative hypothesis), given the null hypothesis ($H_0$) is true.

* **Two-Sided $\mathrm{p}$-Value:** 
$$\displaystyle\mathrm{p}_{\mathrm{obs}} = \mathrm{P}\big(\,|\mathrm{T}|\,\geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big)$$
:::

```{r}
#| echo: false
alpha  <- 0.05
df     <- 5-1
q_1    <- -abs(T_obs)
q_2    <- abs(T_obs)
p_obs  <- round(2 * min(pt(q_1, df = df), pt(q_2, df = df)), 2)
```

The light-red area in @fig-pValue equals the observed $\mathrm{p}$-value [$\mathrm{p}_{\mathrm{obs}}=`r p_obs`$]{style="background-color: #F88379CC; padding: 0.2em 0.1em;"} corresponding to the observed value of the $t$-statistic $\mathrm{T}_{\mathrm{obs}}=`r T_obs`.$

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-pValue
#| fig-cap: "Density of the null-distribution $t_{\\operatorname{df}}$ with $\\operatorname{df}=n-1=4$ (marketing campaign data)."


xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)

par(mar = c(4.5, .1, .1, .1), family = "serif")
curve(expr = dt(x = x, df = df),
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      ylim = c(0, .4), xaxs="i",yaxs="i", axes = FALSE)

polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = myRed)
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = myRed)
curve(expr = dt(x = x, df = df),lwd = 2, col = 1,add = TRUE)
axis(1, cex.axis = 1.5, family = "serif")
axis(1, at =  T_obs, col = "red", labels = FALSE, tck=-0.09)
axis(1, at = -T_obs, col = "red", labels = FALSE, tck=-0.09)
mtext(side = 1, text = expression(-T[obs], T[obs]), at = c(-abs(T_obs),abs(T_obs)), line = 3.5, cex = 1.5, family = "serif")
legend("topleft", pch=c(22), cex=1.5, 
       col = myRed, 
       legend=bquote(p[obs]==.(p_obs)), 
       bty="n", pt.bg = myRed)
```


### Computing the $\mathrm{p}$-Value

Note that 
$$
% \mathrm{T}=\frac{\bar{X}-0}{S/\sqrt{n}}\overset{H_0}{\sim}t_{n-1}
\mathrm{T}=\frac{\bar{X}-0}{S/\sqrt{n}}\overset{H_0}{\sim}t_{n-1}\phantom{Half}
$$
implies that the **absolute value** $|T|$ of the random $t$-statistic is Half-$t_{n-1}$ distributed,
$$
% \mathrm{T}=\frac{\bar{X}-0}{S/\sqrt{n}}\overset{H_0}{\sim}t_{n-1}
|\mathrm{T}|=\frac{|\bar{X}-0|}{S/\sqrt{n}}\overset{H_0}{\sim}\textrm{Half-}t_{n-1}.
$$

@fig-Halft visualizes the density function of the $\textrm{Half-}t_{n-1}$ distribution with $\operatorname{df}=n-1=4$ degrees of freedom for the marketing campaign data ($n=5$).

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-Halft
#| fig-cap: "Densities of the null-distributions $t_{\\operatorname{df}}$ and $\\textrm{Half-}t_{\\operatorname{df}}$ with $\\operatorname{df}=n-1=4$ (marketing campaign data)."

library("extraDistr")

alpha  <- 0.05
df     <- 5-1
q_1    <- -abs(T_obs)
q_2    <- abs(T_obs)

p_obs <- round(2 * min(pt(q_1, df = df), pt(q_2, df = df)), 2)

xx0    <- seq(0.00001, 5, len=100)
##
xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)
##
xx3    <- seq(abs(T_obs),  5, len=25)
yy3    <- dht(x = xx3, nu = df)

par(mar = c(4.5, .1, .5, .1), family = "serif")
plot(y = dht(x = xx0, nu = df), x = xx0,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      xaxs="i",yaxs="i", axes = FALSE, type="l")
polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = gray(.95))
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = gray(.95))
curve(expr = dt(x = x, df = df), lwd = 2, col = gray(.75), add = TRUE)
polygon(x = c(xx3,rev(xx3)), y=c(yy3, rep(0,length(yy3))), border = NA, col = myRed)
lines(x=c(0,0), y = c(0,dht(x = xx0, nu = df)[1]), lwd=2)
lines(y = dht(x = xx0, nu = df), x = xx0, lwd = 2, col = 1)
lines(x=c(0,0), y = c(0,dht(x = xx0, nu = df)[1]), lwd=2)
axis(1, cex.axis = 1.5, family = "serif")
axis(1, at =  T_obs, col = "red", labels = FALSE, tck=-0.11)
#axis(1, at = -T_obs, col = gray(.75), labels = FALSE, tck=-0.11)
mtext(side = 1, text = expression("|"~T[obs]~"|"), at = c(abs(T_obs)), line = 3.5, cex = 1.5, family = "serif")
legend("topleft", pch=22, cex=1.5, 
       col = myRed, 
       legend=c(bquote(p[obs]==.(p_obs))), 
       bty="n", pt.bg = myRed)

legend("topright", lty = 1, lwd=2, col=c(1, gray(.75)),
        legend=c(expression("Half-t"[n-1]), expression("t"[n-1])), 
        bty="n", cex = 1.5, title = "Density", fami)       
```



<!-- 
::: {.callout-tip appearance="minimal"}
**$\mathrm{p}$-Value:** The $\mathrm{p}$-value is the probability, assuming the null hypothesis ($H_0$) is true, of observing a test statistic $\mathrm{T}$ as extreme as $\mathrm{T}_{\mathrm{obs}}$ or more extreme (in the direction(s) of the alternative hypothesis) than $\mathrm{T}_{\mathrm{obs}}$. 
::: 

::: {.callout-tip appearance="minimal"}
Thus for two-sided test statistics, we have
\begin{align*}
\mathrm{p}_{\mathrm{obs}} 
& = \mathrm{P}\big(|\mathrm{T}| \geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big).
\end{align*}
:::
-->

The cumulative distribution function (cdf) of the Half-$t$ distribution allows us to compute the $\mathrm{p}$-value of the two-sided $t$-statistic:

\begin{align*}
\mathrm{p}_{\mathrm{obs}} 
& = \mathrm{P}\big(\,|\mathrm{T}|\,\geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big)\\[1ex]
& = 1 - \mathrm{P}\big(\highlight{|\mathrm{T}|}\leq |\mathrm{T}_{\mathrm{obs}}|\;\big|\highlight{H_0\;\text{is true}}\big)\\[1ex]
& = 1 - \highlight{\mathrm{F}_{\,\text{Half-}t,n-1}}\big(|\mathrm{T}_{\mathrm{obs}}|\big),
\end{align*}
where 
$\mathrm{F}_{\,\text{Half-}t,n-1}$ denotes the cdf of the Half-$t$ distribution with $\operatorname{df}=n-1$


#### **Observed $\mathrm{p}$-Value (Marketing Campaign Data):** {-}

Since $|\mathrm{T}_{\mathrm{obs}}|=`r abs(T_obs)`,$
$$
\mathrm{p}_{\mathrm{obs}}=`r p_obs`=1 - \mathrm{F}_{\,\text{Half-}t,n-1}\big(`r abs(T_obs)`\big).
$$

::: {.panel-tabset}

## Plot

Computation of the observed $\operatorname{p}$-value:

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center

library("extraDistr")
n  <- 5
df <- n - 1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)


alpha  <- 0.05
df     <- 5-1

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5), family = "serif")
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, 1), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.5, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.5, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_obs, col = "blue", labels = FALSE, tck=-0.15)
axis(2, at =  myfun(T_obs), col = "blue", labels = FALSE, tck=-0.05)
lines(x = c(0,T_obs), y= rep(myfun(T_obs), 2), col="blue")
lines(x = rep(T_obs, 2), y= c(0, myfun(T_obs)), col="blue")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"==.(T_obs)), at = c(abs(T_obs)), line = 3.5, cex = 1.5, family = "serif")
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_obs)), bquote(p[obs]==.(round(myfun(T_obs), 2))), srt = 0, cex = 1.5, family = "serif")

legend("top", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```


## `R`-Codes

Computation of the observed $\operatorname{p}$-value:
```{r}
#| echo: true
#| eval: false
library("extraDistr")
n  <- 5
df <- n - 1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)

## Alternatively: 
result <- t.test(X_obs, mu = 0, alternative = "two.sided")
p_obs  <- result$p.value
```
:::


### Test Decision based on the $\mathrm{p}$-Value

* We **reject** $H_0$ and accept $H_1$ at the significance level $0<\alpha<1$ if 
$$
\highlight{\mathrm{p}_{\mathrm{obs}} < \alpha}
$$

* We **fail to reject** $H_0$ at the significance level $0<\alpha<1$ if 
$$
\highlight{\mathrm{p}_{\mathrm{obs}} \geq \alpha}
$$


#### **Test Decision (Marketing Campaign Data):** {-}

Since 
$$
\mathrm{p}_{\mathrm{obs}}=`r round(p_obs,2)` \geq \alpha = 0.05
$$
we **fail to reject** $H_0\colon \mu =0$ at the significance level $\alpha=0.05.$ 

That is, we have no statistical evidence that the marketing campaign showed a tendency toward *significantly* higher or lower sales. 


::: {.callout-caution appearance="minimal"}
⚠️ **Failing to reject $H_0$ does not mean that $H_0$ is true.** It only means that a potential impact of the campaign is not large enough such that the observed effect $(\bar{X}_{\mathrm{obs}}-0)$ was not large enough to stand out from the random variation captured by the standard error $(S_{\mathrm{obs}}/\sqrt{n});$ see @sec-Interpretation.
:::


<!-- 
```{r}
#| echo: true
library("extraDistr")
n  <- 5
df <- n-1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)

qht(1-0.05, nu = df)
``` 
-->


### Equal Test Decisions: Critical Values and $\mathrm{p}$-Values

For a given significance level $0<\alpha<1$, critical values and $\mathrm{p}$-values yield equivalent test decisions:

:::: {.columns}

::: {.column width="50%"}

* Reject $H_0$ 
  * if $|\mathrm{T}_{\mathrm{obs}}| > \mathrm{q}_{1-\alpha}\;$ or
  * if $\,\,\,\mathrm{p}_{\mathrm{obs}}\,<\alpha$

:::

::: {.column width="50%"}

* Fail to reject $H_0$ 
  * if $|\mathrm{T}_{\mathrm{obs}}| \leq  \mathrm{q}_{1-\alpha}\;$ or
  * if $\,\,\,\mathrm{p}_{\mathrm{obs}}\, \leq \alpha$

:::

::::

where $\mathrm{q}_{1-\alpha}$ denotes the $(1-\alpha)$-quantile of the Half-$t_{n-1}$ distribution.

::: {.panel-tabset}

## Reject $H_0$

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center

alpha  <- 0.05
df     <- 5-1

T_2obs <- 3.7

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5))
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, .2), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.5, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.5, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_2obs, col = "blue", labels = FALSE, tck=-0.15)
axis(2, at =  myfun(T_2obs), col = "blue", labels = FALSE, tck=-0.15)
axis(1, at =  qht(1-alpha, nu=df), col = "red", labels = FALSE, tck=-0.15)
axis(2, at = alpha, col = "red", labels = FALSE, tck=-0.15)
lines(x = c(0,T_2obs), y= rep(myfun(T_2obs), 2), col="blue")
lines(x = rep(T_2obs, 2), y= c(0, myfun(T_2obs)), col="blue")
lines(x = c(0,qt(1-alpha/2, df=df)), y= rep(alpha, 2), col="red")
#lines(x = c(0, 0), y= c(0, alpha), lwd = 3, col = "red")
lines(x = rep(qt(1-alpha/2, df=df), 2), y= c(0, alpha), lwd = 1, col = "red")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"), at = c(abs(T_2obs)), line = 3.5, cex = 1.5, family = "serif")
mtext(side = 1, text = bquote(q[1-alpha]), at = qht(1-alpha, nu=df), line = 3.5, cex = 1.5, family = "serif")
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_2obs)), bquote(p[obs]), srt = 0, cex = 1.5, family = "serif")
text(x = corners[1]-.75, y = alpha, bquote(alpha==.(alpha)), srt = 0, cex = 1.5, family = "serif")


legend("topright", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```


## Fail to Reject $H_0$

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center

alpha  <- 0.05
df     <- 5-1

T_2obs <- 2.1

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5))
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, .2), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.5, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.5, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_2obs, col = "blue", labels = FALSE, tck=-0.15)
axis(2, at =  myfun(T_2obs), col = "blue", labels = FALSE, tck=-0.15)
axis(1, at =  qht(1-alpha, nu=df), col = "red", labels = FALSE, tck=-0.15)
axis(2, at = alpha, col = "red", labels = FALSE, tck=-0.15)
lines(x = c(0,T_2obs), y= rep(myfun(T_2obs), 2), col="blue")
lines(x = rep(T_2obs, 2), y= c(0, myfun(T_2obs)), col="blue")
lines(x = c(0,qt(1-alpha/2, df=df)), y= rep(alpha, 2), col="red")
#lines(x = c(0, 0), y= c(0, alpha), lwd = 3, col = "red")
lines(x = rep(qt(1-alpha/2, df=df), 2), y= c(0, alpha), lwd = 1, col = "red")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"), at = c(abs(T_2obs)), line = 3.5, cex = 1.5, family = "serif")
mtext(side = 1, text = bquote(q[1-alpha]), at = qht(1-alpha, nu=df), line = 3.5, cex = 1.5, family = "serif")
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_2obs)), bquote(p[obs]), srt = 0, cex = 1.5, family = "serif")
text(x = corners[1]-.75, y = alpha, bquote(alpha==.(alpha)), srt = 0, cex = 1.5, family = "serif")


legend("topright", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```
:::


### Interpretation of the $\textrm{p}$-Value {#sec-Interpretation}

By its definition, the $\textrm{p}$-value, $\textrm{p}_{\mathrm{obs}},$ tells us the probability of observing a value of $|\mathrm{T}|$ as extreme as or more extreme than $|\mathrm{T}_{\mathrm{obs}}|$ if the null hypothesis were $H_0$.


That is, $\textrm{p}_{\mathrm{obs}}$ measures how extreme $|\mathrm{T}_{\mathrm{obs}}|$ is if the null hypothesis $H_0$ were true. 


::: {.callout-tip appearance="minimal"}

## 🤓 Small $\textrm{p}$-Values

* A very small $\textrm{p}_{\mathrm{obs}}$ suggests that $|\mathrm{T}_{\mathrm{obs}}|$ is incompatible with $H_0$, so we may decide that $H_0$ does not provide a plausible explanation for the data. 
* If we reject $H_0$, since  $\textrm{p}_{\mathrm{obs}}<\alpha,$ we falsely reject $H_0$ (type-I-error) with probability $\alpha,$ which we may find acceptable for small $\alpha$-levels such as $\alpha=0.05$ or $\alpha = 0.01.$ 
:::


::: {.callout-warning appearance="minimal"}

## 😑 Large $\textrm{p}$-Values

A large $\textrm{p}_{\mathrm{obs}}$ suggests that $|\mathrm{T}_{\mathrm{obs}}|$ is compatible with $H_0$, **but this does not allow** us to conclude that $H_0$ is confirmed--- even if $\textrm{p}_{\mathrm{obs}}\approx 1$.

*  Even when $H_0$ is false, we may see very large values of $\textrm{p}_{\mathrm{obs}}$ if, for instance, 
    * the sample size $n$ is small or  
    * the variability $\operatorname{Var}(X)$ in the data is large. 

* If we fail to reject $H_0$, since  $\textrm{p}_{\mathrm{obs}}\geq \alpha,$ we cannot simply accept $H_0$, since we do not know at which probability we falsely accept $H_0$ (type-II-error). 

📢 Statistical hypothesis tests are designed to **falsify** $H_0$ not to verify $H_0.$ 
:::



## $\mathrm{p}$-Values: Challenges for Hypothesis Testing

* While the $\mathrm{p}$-value can be a useful statistical measure, it is commonly misused and misinterpreted. 

::: {.callout-warning appearance="minimal"}

**Don't Do's:**

* $\mathrm{p}$-values do not measure the probability that $H_0$ is true

* $\mathrm{p}$-values do not measure the probability that the data were produced by random chance alone.

* $\mathrm{p}$-values do not measure the size of an effect or the importance of a result.

* Scientific conclusions and business or policy decisions should not be based only on whether a $\mathrm{p}$-value passes a specific $\alpha$-threshold.

See @Wasserstein2016asa
:::


::: {.callout-tip appearance="minimal"}

**Do's:**

* **Be Thoughtful:** Statistically thoughtful researchers ...  
   * recognize when they are doing exploratory studies (for generating statistical hypothesis) or more rigidly pre-planned studies (for testing statistical hypothesis).
   * prioritize sound data production by putting energy into the careful planning, design, and execution
of the study.   

* **Be Open:**

* Be Modest

* Scientific conclusions and business or policy decisions should not be based only on whether a $\mathrm{p}$-value passes a specific $\alpha$-threshold.

See @wasserstein2019asa
:::

* "Scientific method: Statistical errors." @nuzzo2014scientific

* ScienceNews (Siegfried 2010): "It’s science's dirtiest secret: The 'scien-
tific method' of testing hypotheses by statistical analysis stands on a flimsy foundation."


## References {-}

::: {#refs}
:::

<!-- 
## 💡 Tips for Success

<br>

* Don’t just memorize formulas — focus on understanding *why* they work.
* Discuss and collaborate
 -->
