<!-- LTeX: language=en-US -->
# $\mathrm{p}$-Values: Challenges for Hypothesis Testing

::: {.content-hidden unless-format="html"}
$$
\require{color}
%% Colorbox within equation-environments:
\newcommand{\highlight}[2][yellow]{\mathchoice%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  }%
$$
:::

::: {.content-hidden unless-format="pdf"}
```{=latex}
%% Colorbox within equation-environments:
\newcommand{\highlight}[2][yellow]{\mathchoice%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\displaystyle#2$}}%
  }%
```
:::

## Introduction 

### 🎯 Lecture Goals {-}

In this lecture, we’ll tackle three key questions:

* **How** do $\mathrm{p}$-values actually work?
* **Why** can $\mathrm{p}$-values be tricky or even misleading in hypothesis testing?
* **What** is good practice when working with $\mathrm{p}$-values?

We’ll explore these questions using the **two-sided $t$-test** as our main example. Parallel arguments apply to one-sided testing and other hypothesis tests (F-test, etc.)

```{r}
#| echo: false
library("scales")
library("extraDistr")

value <- 0
value <- value + 1
set.seed(33)
n     <- 5
Y_1   <- round(rnorm(n, mean = 5), 2)
Y_2   <- round(rnorm(n, mean = 5), 2)
X     <- round(Y_2 - Y_1, 2)
X_bar <- round(mean(X), 2)
s     <- round(sd(X),   2)
T_obs <- round((sqrt(n) * X_bar)/s, 2)
```


### 💻 Example: Marketing Campaign {-}

Imagine a company runs a marketing campaign in $n = 5$ stores and measures sales *before* and *after* the campaign.

<br>

::: {.panel-tabset}

## Data Table


Store | Sales Before | Sales After | Difference (After $-$ Before)
----|------------|------------|--------- 
$1$   |$`r format(Y_1[1], nsmall=2)`$|$`r format(Y_2[1], nsmall=2)`$|$X_{1,\mathrm{obs}}=`r format(X[1], nsmall=2)`$
$2$   |$`r format(Y_1[2], nsmall=2)`$|$`r format(Y_2[2], nsmall=2)`$|$X_{2,\mathrm{obs}}=`r format(X[2], nsmall=2)`$
$3$   |$`r format(Y_1[3], nsmall=2)`$|$`r format(Y_2[3], nsmall=2)`$|$X_{3,\mathrm{obs}}=`r format(X[3], nsmall=2)`$
$4$   |$`r format(Y_1[4], nsmall=2)`$|$`r format(Y_2[4], nsmall=2)`$|$X_{4,\mathrm{obs}}=`r format(X[4], nsmall=2)`$
$5$   |$`r format(Y_1[5], nsmall=2)`$|$`r format(Y_2[5], nsmall=2)`$|$X_{5,\mathrm{obs}}=`r format(X[5], nsmall=2)`$  

## Data Plot

@fig-diffData visualizes the observed differences $X_{i,\mathrm{obs}},$ $i=1,\dots,5.$

<br>

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 0.75
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-diffData
#| fig-cap: "Sales differences (sales after minus before the marketing campaign)."
par(mar = c(2, 0, 0, 0))
plot(x = X, y = rep(0.002, n), 
     pch = 21, 
     col = alpha("red", 0.5), 
     bg  = alpha("red", 0.5), 
     cex = 2, axes = FALSE, 
     ylab = "", xlab = "", xlim = c(-2.1,2.1), ylim = c(0,.005), yaxs="i")
axis(1, at = -2:2, cex.axis = 1.5, lwd = 2, family="serif", pos =0)
```
:::

**We want to know:**

* **Research question:**  </br> 
Was the marketing campaign successful? Did it show a tendency (location-shift) toward *significantly* higher or lower sales?

* **Statistical question:** </br> 
Is the mean, $\mu,$ of $X_1,\dots,X_n$ different from zero?


::: {.callout-tip appearance="minimal"}
## Assumption
There are no systematic differences (mean-shifts) between sales before and after the marketing campaign, other than the potential impact of the campaign itself.
:::


#### **Two-Sided $t$-Test** {-}
We test

<center>
$\displaystyle H_0\colon \mu = 0\quad$ versus $\displaystyle \quad H_1\colon \mu \neq 0$
</center>

<br>

The $t$-statistic is
$$
\mathrm{T} = \frac{\bar{X} - 0}{S/\sqrt{n}}\overset{H_0}{\sim} t_{n-1},
$$
where 

* $\bar{X}=\frac{1}{n}\sum_{i=1}^n X_i$ is the **sample mean**, and
* $S= \sqrt{\frac{1}{n-1}\sum_{i=1}^n(X_i-\bar{X})}$ is the **sample standard deviation.**

::: {.callout-tip appearance="minimal" collapse="true"}
## Assumptions underlying the $t$-test
* Random (iid) sample: 
$X_1,\dots,X_n\overset{\operatorname{iid}}{\sim}X$ 
* For small $n$: 
$X\sim\mathcal{N}(\mu, \sigma^2)$ 
* For large $n$: we just need finite mean $E(X)=\mu$ and variance $\operatorname{Var}(X)=\sigma^2$ (Central Limit Theorem helps).
:::

<br>

#### Observed (obs) Value of the $t$-statistic<br>(Marketing Example) {-}
$$
\mathrm{T}_{\mathrm{obs}} = \frac{\bar{X}_{\mathrm{obs}} - 0}{S_{\mathrm{obs}}/\sqrt{5}} = `r format(T_obs, nsmall=2)`
$$ 

* $\bar{X}_{\mathrm{obs}}=\frac{1}{5}\sum_{i=1}^5 X_{i,\mathrm{obs}} = `r format(X_bar,nsmall=2)`$ 
* $S_{\mathrm{obs}} = \sqrt{\frac{1}{5-1}\sum_{i=1}^5(X_{i,\mathrm{obs}}-\bar{X}_{\mathrm{obs}})} = `r format(s,nsmall=2)`$ 

::: {.callout-note appearance="minimal" collapse="true"}
## Computing $\mathrm{T}_{\mathrm{obs}}$ in `R`

```{r}
#| echo: true
#| eval: false
n     <- 5
X_obs <- c(0.64, -0.72, -0.23, 0.91, 1.06)
X_bar <- mean(X_obs)
S_obs <- sd(X_obs)

## Observed t-statistic
T_obs <- (X_bar - 0)/(S_obs/sqrt(n))

## Alternatively: 
result <- t.test(x           = X_obs, 
                 mu          = 0, 
                 alternative = "two.sided")
T_obs  <- result$statistic
```
:::

<br>

#### Test Decision using Critical Values<br>(Marketing Example) {-}

Significance level
$$
\alpha=0.05
$$

<br>

Critical values of the $t_{n-1}$-distribution 

<center>
$\mathrm{q}_{\alpha/2} = `r round(qt(p = 0.05/2, df=4), 2)`\;$ and $\;\mathrm{q}_{1-\alpha/2} = `r round(qt(p = 1-0.05/2, df=4), 2)`$
</center>

<br>

Since $\textrm{T}_{\textrm{obs}}$ is not in the rejection regions (see @fig-CriticalValues)
\begin{align*}
\textrm{T}_{\textrm{obs}}=`r T_obs` \;\highlight{\not\in}\;(-\infty, \mathrm{q}_{\alpha/2})\; \cup \;(\mathrm{q}_{1-\alpha/2},\infty\,)
\end{align*}
we [fail to reject $H_0\colon\mu = 0$]{style="background-color: yellow; padding: 0.1em 0.1em;"}  at the significance level [$\alpha = 0.05.$]{style="background-color: #F88379CC; padding: 0.1em 0.1em;"}


```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-CriticalValues
#| fig-cap: "Density of the null-distribution $t_{\\operatorname{df}}$ with $\\operatorname{df}=n-1=4$ (marketing example). The rejection regions are defined by the quantiles (critical values) $\\mathrm{q}_{\\alpha/2}$ and $\\mathrm{q}_{1-\\alpha/2}$ of $t_{\\operatorname{df}}$ for $\\alpha=0.05.$"


alpha  <- 0.05
df     <- 5-1
q_1    <- qt(p = alpha/2,   df=df)
q_2    <- qt(p = 1-alpha/2, df=df)

xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)

myRed  <- "#F88379CC"

par(mar = c(4.5, .1, .1, .1))
curve(expr = dt(x = x, df = df),
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      ylim = c(0, .4), xaxs="i",yaxs="i", axes = FALSE)

polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = myRed)
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = myRed)
curve(expr = dt(x = x, df = df),
      lwd = 2, col = 1, add = TRUE)
axis(1, cex.axis = 1.5, family = "serif")
axis(1, at = T_obs, col = "blue", labels = FALSE, tck=-0.09)
axis(1, at = q_1,   col = "red",  labels = FALSE, tck=-0.09)
lines(x = c(q_1,q_1), y = c(0, dt(x = q_1, df = df)), col="red")
axis(1, at = q_2,   col = "red",  labels = FALSE, tck=-0.09)
lines(x = c(q_2,q_2), y = c(0, dt(x = q_2, df = df)), col="red")
mtext(side = 1, text = c(expression(q[alpha/2]), expression(T[obs]), expression(q[1-alpha/2])), at = c(q_1,T_obs,q_2), line = 3.5, cex = 1.5, family = "serif")
legend("topright", pch=c(22), cex=1.5, 
       col = myRed, 
       legend=expression(alpha==0.05), 
       bty="n", pt.bg = myRed, title = "")
```

::: {.callout-note appearance="minimal" collapse="true"}
## Computing critical values in `R`

```{r}
#| echo: true
#| eval: false
## Critical values
n     <- 5
alpha <- 0.05
qt(p =     alpha/2, df = n-1)
qt(p = 1 - alpha/2, df = n-1)
```
:::



## The Two-Sided $\mathrm{p}$-Value of the $t$-Test 

### Definition of the $\mathrm{p}$-Value

::: {.callout-tip appearance="minimal"}
## Definition of the $\mathrm{p}$-Value 
The $\mathrm{p}$-value, $\textrm{p}_{\mathrm{obs}},$ is the probability of observing a test statistic $\mathrm{T}$ as extreme as or more extreme than $\mathrm{T}_{\mathrm{obs}}$ (in the direction(s) of the alternative hypothesis), given the null hypothesis ($H_0$) is true.
:::

**Two-sided $\mathrm{p}$-value of the $t$-test** 
$$\displaystyle\mathrm{p}_{\mathrm{obs}} = \mathrm{P}\big(\,|\mathrm{T}|\,\geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big)$$


<br>

```{r}
#| echo: false
alpha  <- 0.05
df     <- 5-1
q_1    <- -abs(T_obs)
q_2    <- abs(T_obs)
p_obs  <- round(2 * min(pt(q_1, df = df), pt(q_2, df = df)), 2)
```

The light-red area in @fig-pValue equals the observed $\mathrm{p}$-value [$\mathrm{p}_{\mathrm{obs}}=`r p_obs`$]{style="background-color: #F88379CC; padding: 0.2em 0.1em;"} corresponding to the observed value of the $t$-statistic $\mathrm{T}_{\mathrm{obs}}=`r T_obs`.$

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-pValue
#| fig-cap: "Density of the null-distribution $t_{\\operatorname{df}}$ with $\\operatorname{df}=n-1=4$ (marketing example)."


xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)

par(mar = c(4.5, .1, .1, .1), family = "serif")
curve(expr = dt(x = x, df = df),
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      ylim = c(0, .4), xaxs="i",yaxs="i", axes = FALSE)

polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = myRed)
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = myRed)
curve(expr = dt(x = x, df = df),lwd = 2, col = 1,add = TRUE)
axis(1, cex.axis = 1.5, family = "serif")
axis(1, at =  T_obs, col = "red", labels = FALSE, tck=-0.09)
axis(1, at = -T_obs, col = "red", labels = FALSE, tck=-0.09)
mtext(side = 1, text = expression(-T[obs], T[obs]), at = c(-abs(T_obs),abs(T_obs)), line = 3.5, cex = 1.5, family = "serif")
legend("topright", pch=c(22), cex=1.5, 
       col = myRed, 
       legend=bquote(p[obs]==.(p_obs)), 
       bty="n", pt.bg = myRed, title = "")
```


### Computing the $\mathrm{p}$-Value

Note that 
$$
% \mathrm{T}=\frac{\bar{X}-0}{S/\sqrt{n}}\overset{H_0}{\sim}t_{n-1}
\mathrm{T}=\frac{\bar{X}-0}{S/\sqrt{n}}\overset{H_0}{\sim}t_{n-1}\phantom{Half}
$$
implies that the [**absolute value** $|\mathrm{T}|$]{style="background-color: yellow; padding: 0.1em 0.1em;"} of the random $t$-statistic is Half-$t_{n-1}$ distributed,
$$
% \mathrm{T}=\frac{\bar{X}-0}{S/\sqrt{n}}\overset{H_0}{\sim}t_{n-1}
|\mathrm{T}|=\frac{|\bar{X}-0|}{S/\sqrt{n}}\overset{H_0}{\sim}\highlight{\textrm{Half-}t_{n-1}.}
$$

<br>

The density function of the $\textrm{Half-}t_{n-1}$ distribution with $\operatorname{df}=n-1$ degrees of freedom is simply the **folded version** of the density function of the $t_{n-1}$ distribution (@fig-Halft).

<br>

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center
#| label: fig-Halft
#| fig-cap: "Densities of the null-distributions $t_{\\operatorname{df}}$ and $\\textrm{Half-}t_{\\operatorname{df}}$ with $\\operatorname{df}=n-1=4$ (marketing example)."

library("extraDistr")

alpha  <- 0.05
df     <- 5-1
q_1    <- -abs(T_obs)
q_2    <- abs(T_obs)

p_obs <- round(2 * min(pt(q_1, df = df), pt(q_2, df = df)), 2)

xx0    <- seq(0.00001, 5, len=100)
##
xx1    <- seq(-5, q_1, len=25)
yy1    <- dt(x = xx1, df = df)
##
xx2    <- seq(q_2,  5, len=25)
yy2    <- dt(x = xx2, df = df)
##
xx3    <- seq(abs(T_obs),  5, len=25)
yy3    <- dht(x = xx3, nu = df)

par(mar = c(4.5, .1, .5, .1), family = "serif")
plot(y = dht(x = xx0, nu = df), x = xx0,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(-5, 5),
      xaxs="i",yaxs="i", axes = FALSE, type="l")
polygon(x = c(xx1,rev(xx1)), y=c(yy1, rep(0,length(yy1))), border = NA, col = gray(.95))
polygon(x = c(xx2,rev(xx2)), y=c(yy2, rep(0,length(yy2))), border = NA, col = gray(.95))
curve(expr = dt(x = x, df = df), lwd = 2, col = gray(.75), add = TRUE)
polygon(x = c(xx3,rev(xx3)), y=c(yy3, rep(0,length(yy3))), border = NA, col = myRed)
lines(x=c(0,0), y = c(0,dht(x = xx0, nu = df)[1]), lwd=2)
lines(y = dht(x = xx0, nu = df), x = xx0, lwd = 2, col = 1)
lines(x=c(0,0), y = c(0,dht(x = xx0, nu = df)[1]), lwd=2)
axis(1, cex.axis = 1.5, family = "serif")
axis(1, at =  T_obs, col = "red", labels = FALSE, tck=-0.11)
#axis(1, at = -T_obs, col = gray(.75), labels = FALSE, tck=-0.11)
mtext(side = 1, text = expression("|"~T[obs]~"|"), at = c(abs(T_obs)), line = 3.5, cex = 1.5, family = "serif")
legend("topright", pch=22, cex=1.5, 
       col = myRed, 
       legend=c(bquote(p[obs]==.(p_obs))), 
       bty="n", pt.bg = myRed, title = "")

legend("topleft", lty = 1, lwd=2, col=c(1, gray(.75)),
        legend=c(expression(paste("Half-",italic(t)[n-1])), expression(italic(t)[n-1])), 
        bty="n", cex = 1.5, title = "Density")       
```

::: {.callout-tip appearance="minimal"}
## Computing the $\mathrm{p}$-Value
<br>
The cumulative distribution function of the Half-$t$ distribution, 
$$
\mathrm{F}_{\,\text{Half-}t,n-1}(x)=\mathrm{P}(|\textrm{T}|\leq x),\quad |\textrm{T}|\sim \textrm{Half-}t_{n-1}
$$ 
allows us to compute the $\mathrm{p}$-value, $\mathrm{p}_{\mathrm{obs}},$ for the two-sided $t$-statistic $|\mathrm{T}_{\mathrm{obs}}|$:

\begin{align*}
\mathrm{p}_{\mathrm{obs}} 
& = \mathrm{P}\big(\,|\mathrm{T}|\,\geq |\mathrm{T}_{\mathrm{obs}}|\;\big|\;H_0\;\text{is true}\big)\\[1ex]
& = 1 - \mathrm{P}\big(\highlight{|\mathrm{T}|}\leq |\mathrm{T}_{\mathrm{obs}}|\;\big|\highlight{H_0\;\text{is true}}\big)\\[1ex]
& = 1 - \highlight{\mathrm{F}_{\,\text{Half-}t,n-1}}\big(|\mathrm{T}_{\mathrm{obs}}|\big)
\end{align*}
:::

#### Computing the $\mathrm{p}$-Value<br>(Marketing Example) {-}

<br>

\begin{align*}
\mathrm{p}_{\mathrm{obs}}
&=1 - \mathrm{F}_{\,\text{Half-}t,n-1}\big(|\mathrm{T}_{\mathrm{obs}}|\big)\\
&=1 - \mathrm{F}_{\,\text{Half-}t,n-1}\big(\,`r abs(T_obs)`\,\big)=`r p_obs`
\end{align*}

<br>

::: {.panel-tabset}

## Plot

Computation of the observed $\operatorname{p}$-value:

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center

library("extraDistr")
n  <- 5
df <- n - 1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)


alpha  <- 0.05
df     <- 5-1

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5), family = "serif")
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, 1), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.5, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.5, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_obs, col = "blue", labels = FALSE, tck=-0.15)
axis(2, at =  myfun(T_obs), col = "blue", labels = FALSE, tck=-0.05)
lines(x = c(0,T_obs), y= rep(myfun(T_obs), 2), col="blue")
lines(x = rep(T_obs, 2), y= c(0, myfun(T_obs)), col="blue")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"==.(T_obs)), at = c(abs(T_obs)), line = 3.5, cex = 1.5, family = "serif")
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_obs)), bquote(p[obs]==.(round(myfun(T_obs), 2))), srt = 0, cex = 1.5, family = "serif")

legend("top", lty = 1, lwd=2,
       legend=expression(1-F[paste("Half-",italic(t),",n-1")]), 
       bty="n", cex = 1.5)
```


## `R`-Codes

Computation of the observed $\operatorname{p}$-value:
```{r}
#| echo: true
#| eval: false
library("extraDistr")
n  <- 5
df <- n - 1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)

## Alternatively: 
result <- t.test(X_obs, mu = 0, alternative = "two.sided")
p_obs  <- result$p.value
```
:::

<br>

### Test Decision based on the $\mathrm{p}$-Value

<br>

* We **reject** $H_0$ at the significance level $0<\alpha<1$ if 
$$
\highlight{\mathrm{p}_{\mathrm{obs}} < \alpha}
$$

* We **fail to reject** $H_0$ at the significance level $0<\alpha<1$ if 
$$
\highlight{\mathrm{p}_{\mathrm{obs}} \geq \alpha}
$$


#### Test Decision (Marketing Example) {-}

Since 
$$
\mathrm{p}_{\mathrm{obs}}=`r round(p_obs,2)` \geq \alpha = 0.05
$$
we **fail to reject** $H_0\colon \mu =0$ at the significance level $\alpha=0.05.$ 

> We also fail to reject $H_0$ at any other typically used significance level such as, for instance, $\alpha=0.1$ or $\alpha=0.01.$ 

<!-- That is, we have no statistical evidence that the marketing campaign showed a tendency toward *significantly* higher or lower sales.  -->


::: {.callout-caution icon="false" appearance="minimal"}
## ⚠️ Failing to reject $H_0$ does not mean that $H_0$ is true.
<br>
Failing to reject $H_0$ only means that a potential impact of the marketing campaign is not large enough to stand out from the random variation captured by the standard error $(S_{\mathrm{obs}}/\sqrt{n});$ see @sec-Interpretation.
:::


<!-- 
```{r}
#| echo: true
library("extraDistr")
n  <- 5
df <- n-1

## 1 minus the distribution function of the Half-t
p_obs <- 1 - pht(q = abs(T_obs), nu = df)

qht(1-0.05, nu = df)
``` 
-->

<br>

### $\mathrm{p}$-Values versus Critical Values

$\mathrm{p}$-values and critical values lead to exactly **equivalent test decisions:**

:::: {.columns}

::: {.column width="50%"}

* Reject $H_0$ 
  * if $|\mathrm{T}_{\mathrm{obs}}| > \mathrm{q}_{1-\alpha}\;$ or
  * if $\,\,\,\mathrm{p}_{\mathrm{obs}}\,<\alpha$

:::

::: {.column width="50%"}

* Fail to reject $H_0$ 
  * if $|\mathrm{T}_{\mathrm{obs}}| \leq  \mathrm{q}_{1-\alpha}\;$ or
  * if $\,\,\,\mathrm{p}_{\mathrm{obs}}\, \leq \alpha$

:::

::::

where $\mathrm{q}_{1-\alpha}$ denotes the $(1-\alpha)$-quantile of the Half-$t_{n-1}$ distribution.

::: {.panel-tabset}

## Reject $H_0$

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center

alpha  <- 0.05
df     <- 5-1

T_2obs <- 3.7

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5))
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, .2), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.5, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.5, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_2obs, col = "blue", labels = FALSE, tck=-0.15)
axis(2, at =  myfun(T_2obs), col = "blue", labels = FALSE, tck=-0.15)
axis(1, at =  qht(1-alpha, nu=df), col = "red", labels = FALSE, tck=-0.15)
axis(2, at = alpha, col = "red", labels = FALSE, tck=-0.15)
lines(x = c(0,T_2obs), y= rep(myfun(T_2obs), 2), col="blue")
lines(x = rep(T_2obs, 2), y= c(0, myfun(T_2obs)), col="blue")
lines(x = c(0,qt(1-alpha/2, df=df)), y= rep(alpha, 2), col="red")
#lines(x = c(0, 0), y= c(0, alpha), lwd = 3, col = "red")
lines(x = rep(qt(1-alpha/2, df=df), 2), y= c(0, alpha), lwd = 1, col = "red")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"), at = c(abs(T_2obs)), line = 3.5, cex = 1.5, family = "serif")
mtext(side = 1, text = bquote(q[1-alpha]), at = qht(1-alpha, nu=df), line = 3.5, cex = 1.5, family = "serif")
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_2obs)), bquote(p[obs]), srt = 0, cex = 1.5, family = "serif")
text(x = corners[1]-.75, y = alpha, bquote(alpha==.(alpha)), srt = 0, cex = 1.5, family = "serif")


legend("topright", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```


## Fail to Reject $H_0$

```{r}
#| echo: false
#| fig.width: 8
#| fig.height: 4
#| out.width: "1\\textwidth"
#| fig-align: center

alpha  <- 0.05
df     <- 5-1

T_2obs <- 2.1

myfun <- function(x){1-pht(q = x, nu = df)}

par(mar = c(4.5, 9, 1, 0.5))
curve(expr = myfun,
      xlab = "", ylab = "", main = "",
      lwd = 2, col = 1, xlim = c(0, 4),
      ylim = c(0, .2), xaxs="i",yaxs="i", axes = FALSE)
axis(1, cex.axis = 1.5, family = "serif", mgp = c(3, 1, 0))
axis(2, at = c(0,1), labels = c("0","1"), cex.axis = 1.5, family = "serif", las=2, mgp = c(3, 1, 0))
axis(1, at =  T_2obs, col = "blue", labels = FALSE, tck=-0.15)
axis(2, at =  myfun(T_2obs), col = "blue", labels = FALSE, tck=-0.15)
axis(1, at =  qht(1-alpha, nu=df), col = "red", labels = FALSE, tck=-0.15)
axis(2, at = alpha, col = "red", labels = FALSE, tck=-0.15)
lines(x = c(0,T_2obs), y= rep(myfun(T_2obs), 2), col="blue")
lines(x = rep(T_2obs, 2), y= c(0, myfun(T_2obs)), col="blue")
lines(x = c(0,qt(1-alpha/2, df=df)), y= rep(alpha, 2), col="red")
#lines(x = c(0, 0), y= c(0, alpha), lwd = 3, col = "red")
lines(x = rep(qt(1-alpha/2, df=df), 2), y= c(0, alpha), lwd = 1, col = "red")
##
mtext(side = 1, text = bquote("|"~T[obs]~"|"), at = c(abs(T_2obs)), line = 3.5, cex = 1.5, family = "serif")
mtext(side = 1, text = bquote(q[1-alpha]), at = qht(1-alpha, nu=df), line = 3.5, cex = 1.5, family = "serif")
corners = par("usr") #Gets the four corners of plot area (x1, x2, y1, y2)
par(xpd = TRUE) #Draw outside plot area
text(x = corners[1]-.75, y = mean(myfun(T_2obs)), bquote(p[obs]), srt = 0, cex = 1.5, family = "serif")
text(x = corners[1]-.75, y = alpha, bquote(alpha==.(alpha)), srt = 0, cex = 1.5, family = "serif")


legend("topright", lty = 1, lwd=2,
       legend=expression(1-F["Half-t,n-1"]), 
       bty="n", cex = 1.5)
```
:::


## $\mathrm{p}$-Values: Challenges for Hypothesis Testing

$\mathrm{p}$-values can be helpful tools, but they are often **misused and misunderstood**, leading to a large amount of published **“bad science.”**  
This has sparked an ongoing debate about whether $\mathrm{p}$-values — and the idea of **statistical significance** — should be used at all.

* [The problem with p Values: How significant are they, really?](https://sciencex.com/wire-news/145707973/the-problem-with-p-values-how-significant-are-they-really.html) [@Cumming_2013]
* [Scientific method: Statistical errors](https://www.nature.com/articles/506150a) [@Nuzzo_2014]
* [Scientists rise up against statistical significance](https://www.nature.com/articles/d41586-019-00857-9) [@Amrhein_2019]

To address this, the **American Statistical Association (ASA)** released two statements explaining **what $\mathrm{p}$-values and statistical significance can — and cannot — tell us.**

* [The ASA statement on p-values: Context, process, and purpose](https://www.tandfonline.com/doi/full/10.1080/00031305.2016.1154108#d1e167) [@Wasserstein2016asa]
* [Moving to a world beyond "p< 0.05"](https://www.tandfonline.com/doi/full/10.1080/00031305.2019.1583913) [@wasserstein2019asa]

### Correct Interpretation of the $\textrm{p}$-Value {#sec-Interpretation}

The $\textrm{p}$-value $\textrm{p}_{\mathrm{obs}}$ measures how extreme $|\mathrm{T}_{\mathrm{obs}}|$ is if the null hypothesis $H_0$ were true. 

<!-- ::: {.callout-tip appearance="minimal"}

## 🤓 Small $\textrm{p}$-Values
<br>

* A very small $\textrm{p}_{\mathrm{obs}}$ suggests that $|\mathrm{T}_{\mathrm{obs}}|$ is **incompatible** with $H_0.$ So we may decide that $H_0$ does not provide a plausible explanation for the data. 
* If we reject $H_0$, since  $\textrm{p}_{\mathrm{obs}}<\alpha,$ we falsely reject $H_0$ (type-I-error) with probability $\alpha,$ which we may find acceptable for small $\alpha$-levels such as $\alpha=0.05$ or $\alpha = 0.01.$ 
::: -->

::: {.callout-tip appearance="minimal" collapse="true"}

## 🤓 Small $\textrm{p}$-Values: $\;\textrm{p}_{\mathrm{obs}} < \alpha$<br>

* A small $\textrm{p}_{\mathrm{obs}}$ suggests that the observed statistic $|\mathrm{T}_{\mathrm{obs}}|$ is **inconsistent with** the null hypothesis $H_0$.  
  In that case, we may decide that $H_0$ **does not provide a plausible explanation** for the data.

* When we reject $H_0$ (because $\textrm{p}_{\mathrm{obs}} < \alpha$), we accept a risk of making a **type I error** — that is, **falsely rejecting** a true $H_0$ — with probability $\alpha$.  
  This risk is typically considered acceptable for small significance levels such as $\alpha = 0.05$ or $\alpha = 0.01$.

:::

<!-- 
::: {.callout-warning appearance="minimal"}

## 😑 Large $\textrm{p}$-Values
<br>

* A large $\textrm{p}_{\mathrm{obs}}$ suggests that $|\mathrm{T}_{\mathrm{obs}}|$ is **compatible** with $H_0.$ However, this does **not allow** us to conclude that $H_0$ is confirmed. Even very large $p$-values $\textrm{p}_{\mathrm{obs}}\approx 1$ occur simply because a given violation of $H_0$ is smaller than the involved estimation errors; namely if, for instance, 
    * the sample size $n$ is small or  
    * the variability $\operatorname{Var}(X)$ in the data is large. 

* If we fail to reject $H_0$, since  $\textrm{p}_{\mathrm{obs}}\geq \alpha,$ we cannot simply accept $H_0$, since we do not know at which probability we falsely accept $H_0$ (type-II-error). 

<center>
📢 Statistical hypothesis tests are designed to **falsify** $H_0$ not to verify $H_0.$ 
</center>
:::
 -->
::: {.callout-warning appearance="minimal" collapse="true"}

## ⚠️ Large $\textrm{p}$-Values: $\;\textrm{p}_{\mathrm{obs}} \geq \alpha$<br>

* A large $\textrm{p}_{\mathrm{obs}}$ means that the observed test statistic $|\mathrm{T}_{\mathrm{obs}}|$ is **consistent with** the null hypothesis $H_0$.  
  ❗️ However, this does **not** mean that $H_0$ is *true* or *confirmed*.  

* Even very large $\textrm{p}$-values (e.g., $\textrm{p}_{\mathrm{obs}} \approx 1$) can occur simply because a given deviation from $H_0$ does not stand out from the random variation captured by the standard error, for example when  
  * the sample size $n$ is small, or  
  * the variability $\operatorname{Var}(X)$ in the data is large.  

* When we fail to reject $H_0$ (because $\textrm{p}_{\mathrm{obs}} \geq \alpha$), we **cannot accept $H_0$** as we don’t know the probability of making a **type II error** (falsely accepting $H_0$).

<!-- <center>
📢 Statistical hypothesis tests are designed to **falsify** $H_0$ not to verify $H_0.$ 
</center> -->
<center>
📢 Statistical tests are designed to **reject** $H_0$ when there is enough evidence — not to **prove** that $H_0$ is true.
</center>
:::

### General Don't Do's and Do's 

::: {.callout-warning appearance="minimal" collapse="true"}

## ⚠️ Don't Do's<br>

* $\mathrm{p}$-values do not measure the probability that $H_0$ is true.

* $\mathrm{p}$-values do not measure the probability that the data were produced by random chance alone.

* $\mathrm{p}$-values do not measure the size of an effect or the importance of a result.

* Scientific conclusions and business or policy decisions should not be based only on whether a $\mathrm{p}$-value passes a specific $\alpha$-threshold.
:::


::: {.callout-tip appearance="minimal" collapse="true"}

## 🤓 Do's<br> 

* **Be Thoughtful:** Statistically thoughtful researchers ...  
   * recognize when they are doing exploratory studies (for generating statistical hypothesis) or more rigidly pre-planned studies (for testing statistical hypothesis).
   * prioritize sound data production by putting energy into the careful planning, design, and execution
of the study.   

* **Be Open:**
   * Openness to transparency (share data, make results reproducible, etc.)
   * Openness in communication (statistical significance does not mean scientific importance, one study is rarely enough, etc.) 

* **Be Modest:**
   * There is no true statistical/econometric model.
   * Be modest about the role of statistical inference in scientific inference.

<center>
📢 Accept that there will always be uncertainty, and be thoughtful, open, and modest.
</center>
:::

## References {-}

::: {#refs}
:::

